---
title: "The Final Project"
author: "Federico Vicentini, Alice Pratesi, Riccardo Dal Cero, Xhesjana Shametaj"
date: "`r Sys.Date()`"
output: html_document
---

RICORDA DI FARE UNA INTRODUZIONE SUL PAPER

```{r c1, message = FALSE, warning = FALSE}
rm(list = ls())

# Set the working directory to source file location
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Install packages
packages <- c("tidyverse", "rsdmx", "eurostat", "tbl2xts",
              "tidyquant", "BCDating", "pwt10", "dplyr",
              "stargazer", "car", "forecast", "tseries",
              "quantmod", "eurostat", "stargazer",
              "skedastic","Metrics","mFilter", "aTSA","lmtest","xts", "prediction")
new.packages <- packages[!(packages %in% installed.packages()[, "Package"])]
if (length(new.packages)) install.packages(new.packages)
invisible(lapply(packages, library, character.only = TRUE))

# Load packages
library(quantmod)
library(eurostat)
library(FKF)
library(arfima)
```


First we imported the data from the Fred database checking for NA and NULL values
```{r c2, message = FALSE, warning = FALSE}
# Import data
# Download inflation rates from Fred
freddata <- c("PCEPI")
for (i in 1:length(freddata)) {
  getSymbols(freddata[i], src = "FRED")
}
plot(PCEPI)

#Check for NA values
nas=c(which(is.na(PCEPI)))
nas

#Check for null values
zeroes=c(which(PCEPI==0))
zeroes
```


We now use two for cycles to transform PCEPI in PCE inflation rates which is a time series with monthly year on year inflation rates. 
Obviously, by doing this we loose the first twelve observations thus, we remove them from the time series. Also, we cut the time series starting from January 1993.

```{r c3, message = FALSE, warning = FALSE}
inflationus=c()
for(i in 13:length(PCEPI)){
  inflationus[i]=((as.numeric(PCEPI[i])-as.numeric(PCEPI[i-12]))/(as.numeric(PCEPI[i-12])))*100
}
plot(inflationus, type="l")
for(i in 13:length(PCEPI)){
  PCEPI[i]=inflationus[i]
}
PCEPI[1:12]=NA
PCEPI=na.omit(PCEPI)
PCEPI=PCEPI["1993/2022-01-01"]

plot(PCEPI, type="l")
```


Inflation data for Canada was unfortunately available in unpleasant format[:(] thus, we needed to filter only "All-items" values and only for "Canada" as a whole.
Then, we used the same for cycles as before to convert the time series to inflation rates.
```{r c4, message = FALSE, warning = FALSE}
#Canada
canadacpi=read.csv("CanadaCPI3.csv", sep=",")
canadacpi <- canadacpi %>%
  filter(Products.and.product.groups == "All-items") %>%
  filter(GEO == "Canada")
cutcanadacpi=as.data.frame(canadacpi[,11])
cutcanadacpi[,2]=canadacpi[,1]
cutcanadacpi[,3]=cutcanadacpi[,2]
cutcanadacpi[,2]=cutcanadacpi[,1]
cutcanadacpi[,1]=cutcanadacpi[,3]
cutcanadacpi[,3]=NULL
names(cutcanadacpi)=c("Date","Rate")
cpican=cutcanadacpi


inflationcan=c()
for(i in 13:length(cpican$Rate)){
  inflationcan[i]=((as.numeric(cpican$Rate[i])-as.numeric(cpican$Rate[(i-12)]))/(as.numeric(cpican$Rate[(i-12)])))*100
}
plot(inflationcan, type="l")
for(i in 13:length(cpican$Rate)){
  cpican$Rate[i]=inflationcan[i]
}
cpican[1:12,]=NA
cpican=na.omit(cpican)
cpican=cpican[25:length(cpican$Rate),]

plot(cpican$Rate, type="l")
```



Data for Norway was presented in an even worse format than Canadian data. In fact, every row represented a different year and every column was filled with a different month. Thus, we had to design a for cycle in order to straighten up the data in an orderly manner. 
Then, we used the same for cycles as before to convert the time series to inflation rates.
```{r c5, message = FALSE, warning = FALSE}
#Norway
norwaycpi=read.csv("Norway-CPI.csv", sep=";")
norwaydata=matrix(NA,1,2)
norwaydata=as.data.frame(norwaydata)
names(norwaydata)=c("Date","Rate")
count=1
for (i in 2:length(norwaycpi$X)){
  for (s in 12:1){
    norwaydata[count,2]=norwaycpi[i,2+s]
    count=count+1
  }
}
norwaydata$Rate=rev(norwaydata$Rate)
plot(norwaydata$Rate)
norwaydata$Date <- data.frame(time = seq(as.Date('1929-01-01'), by = 'months', length = 1116))

inflationnor=c()
for(i in 13:length(norwaydata$Rate)){
  inflationnor[i]=((as.numeric(norwaydata$Rate[i])-as.numeric(norwaydata$Rate[(i-12)]))/(as.numeric(norwaydata$Rate[(i-12)])))*100
}
plot(inflationnor, type="l")
for(i in 13:length(norwaydata$Rate)){
  norwaydata$Rate[i]=inflationnor[i]
}
norwaydata[1:12,]=NA
norwaydata=na.omit(norwaydata)
norwaydata=norwaydata[757:1116,]

plot(norwaydata$Date,norwaydata$Rate, type="l")
```


Data for Sweden was unexpectedly well presented and also it was already converted into inflation rates. 
```{r c6, message = FALSE, warning = FALSE}
#Sweden
swedencpi=read.csv("Sweden-CPI.csv", sep=" ")
names(swedencpi)=c("Date","Rate")
swedencpi=swedencpi[73:length(swedencpi$Date),]
```

Data for UK was also well presented and already converted into inflation rates. 
```{r c7, message = FALSE, warning = FALSE}
#UK
ukcpi=read.csv("UK-CPI.csv", sep=",")
ukcpi=ukcpi[222:length(ukcpi$Title),]
names(ukcpi)=c("Date","Rate")
```

Data for EU was also well presented and already converted into inflation rates. 
```{r c8, message = FALSE, warning = FALSE}
#EU (from 1999 onward)
eucpi=read.csv("eu-CPI.csv", sep=",")
eucpi=eucpi[4466:4770,7:8]
eucpi=eucpi[25:length(eucpi$TIME_PERIOD),]
names(eucpi)=c("Date","Rate")
```


Data for Germany was well presented but nonetheless we had to convert it into inflation rates using the usual two for cycles.
```{r c9, message = FALSE, warning = FALSE}
#Germany (from 1993 onward)
gercpi=read.csv("Germany-CPI.csv", sep=",")
gercpi=gercpi[,c(1,2)]
names(gercpi)=c("Date","Rate")

inflationger=c()
for(i in 13:length(gercpi$Rate)){
  inflationger[i]=((as.numeric(gercpi$Rate[i])-as.numeric(gercpi$Rate[(i-12)]))/(as.numeric(gercpi$Rate[(i-12)])))*100
}
plot(inflationger, type="l")
for(i in 13:length(gercpi$Rate)){
  gercpi$Rate[i]=inflationger[i]
}
gercpi[1:12,]=NA
gercpi=na.omit(gercpi)
```



```{r c10, message = FALSE, warning= FALSE}
#Graph with all the inflation rates

plot(cpican$Rate, type="l", col="blue", dev="svg")
lines(norwaydata$Rate, type="l", col="red")
lines(gercpi$Rate, type="l", col="green")
eucpigraph=eucpi
nas=matrix(NA,72,1)
eucpigraph=append(eucpigraph$Rate, nas,after=0)
lines(eucpigraph, type="l", col="yellow")
lines(ukcpi$Rate, type="l", col="pink")
lines(swedencpi$Rate, type="l", col="grey")
lines(PCEPI)
```


#Divide time series in chunks for analysis

```{r c11, message= FALSE, warning= FALSE}
#Euro area
eupart1=data.frame(eucpi[1:228,])
eupart2=data.frame(eucpi[1:96,])

#Germany
gerpart1=data.frame(gercpi[1:300,])
gerpart2=data.frame(gercpi[1:168,])
gerpart3=data.frame(gercpi[73:300,])

#UK
ukpart1=data.frame(ukcpi[1:300,])
ukpart2=data.frame(ukcpi[1:168,])
ukpart3=data.frame(ukcpi[73:300,])

#US
uscpi=PCEPI
uspart1=data.frame(uscpi[1:300,])
uspart2=data.frame(uscpi[1:168,])
uspart3=data.frame(uscpi[73:300,])

#Canada
cancpi=cpican
canpart1=data.frame(cancpi[1:300,])
canpart2=data.frame(cancpi[1:168,])
canpart3=data.frame(cancpi[73:300,])

#Norway
norcpi=norwaydata
norpart1=data.frame(norcpi[1:300,])
norpart2=data.frame(norcpi[1:168,])
norpart3=data.frame(norcpi[73:300,])

#Sweden
swecpi=swedencpi
swepart1=data.frame(swecpi[1:300,])
swepart2=data.frame(swecpi[1:168,])
swepart3=data.frame(swecpi[73:300,])
```


```{r c12, message= FALSE, warning= FALSE}

```

